---
title: "Medaka_RNA_seq_analysis"
output:
  html_document: default
  pdf_document: default
date: "2026-06-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!requireNamespace("tidyverse", quietly = TRUE)){install.packages("tidyverse")}
if(!requireNamespace("stringr", quietly = TRUE)){install.packages("stringr")}
if(!requireNamespace("gplots", quietly = TRUE)){install.packages("gplots")}
if(!requireNamespace("matrixStats", quietly = TRUE)){install.packages("matrixStats")}
if(!requireNamespace("extrafont", quietly = TRUE)){install.packages("extrafont")}
if(!requireNamespace("ggpubr", quietly = TRUE)){install.packages("ggpubr")}
if(!requireNamespace("gt", quietly = TRUE)){install.packages("gt")}
if (!requireNamespace("readxl", quietly = TRUE)) {install.packages("readxl")}
if (!requireNamespace("curl", quietly = TRUE)) {install.packages("curl")}
if (!requireNamespace("ggVennDiagram", quietly = TRUE)) { install.packages("ggVennDiagram") }
if (!require("pbapply")) install.packages('pbapply')
library(reticulate)
library(pbapply)
library(tidyverse)
library(stringr)
library(gplots)
library(matrixStats)
library(extrafont)
library(ggpubr)
library(gt)
theme_set(theme_classic(base_size = 6, base_family = "Arial"))
font_import(pattern = "Arial")
loadfonts(device = "all")
```

# Comparison of outdoor transcriptome and signal importance.

### Loading outdoor RNA-seq data

```{r Loading}

#Construct data set
maindir = "../data/RNAseq_nakayama_2015to17"

# Data file lists in Nakayama et al 2023
rep1_file_list <- list.files(file.path(maindir,"group1"))[-1] # Exclude 150929 data 
rep2_file_list <- list.files(file.path(maindir,"group2"))[-1] # Exclude 150929 data 
gene_profiles <- read.csv(file.path(maindir, "group1", rep1_file_list[1])) %>% dplyr::select(-expected_count, -FPKM)

# Construct count matrix from data file
Create_time_corse_rna_seq_df <- function(group, fpkm_file_list){
  for( filename in fpkm_file_list){
    extracted_txt = str_extract(filename, "\\d{6}")
    date = as.Date(extracted_txt, format = "%y%m%d")
    formatted_date <- format(date, "%y-%m/%d")
    if(filename ==  fpkm_file_list[1]){
      RNAseq_time_series_df <- read.csv(file.path(maindir, group, filename)) %>% dplyr::select(gene_id, Symbol, FPKM) 
    }else{
      data <- read.csv(file.path(maindir, group, filename)) %>% dplyr::select(gene_id, FPKM)
      RNAseq_time_series_df <- right_join(RNAseq_time_series_df, data, by="gene_id")
    }
    colnames(RNAseq_time_series_df)[colnames(RNAseq_time_series_df) == "FPKM"] <- formatted_date
  }
  return(RNAseq_time_series_df)
}

rep1_RNAseq_time_series_df = Create_time_corse_rna_seq_df("group1", rep1_file_list)
rep2_RNAseq_time_series_df = Create_time_corse_rna_seq_df("group2", rep2_file_list)
symbol_to_geneid <- rep1_RNAseq_time_series_df %>% dplyr::select(gene_id, Symbol)

#Dimension is different due to quality control. Taking average in the shared genes.
shared_genes <- intersect(rep1_RNAseq_time_series_df$gene_id, rep2_RNAseq_time_series_df$gene_id) 
rep1_matrix <- rep1_RNAseq_time_series_df %>% arrange(gene_id) %>% column_to_rownames("gene_id") %>% dplyr::select(-Symbol)
rep2_matrix <- rep2_RNAseq_time_series_df %>% arrange(gene_id) %>% column_to_rownames("gene_id") %>% dplyr::select(-Symbol)

rep1_shared_gene_matrix <- rep1_RNAseq_time_series_df %>% filter(is.element(gene_id, shared_genes)) %>% arrange(gene_id) %>% column_to_rownames("gene_id") %>% dplyr::select(-Symbol)
rep2_shared_gene_matrix <- rep2_RNAseq_time_series_df %>% filter(is.element(gene_id, shared_genes)) %>% arrange(gene_id) %>% column_to_rownames("gene_id") %>% dplyr::select(-Symbol)
is.order.same <-  all(rownames(rep1_shared_gene_matrix)==rownames(rep2_shared_gene_matrix))
if(is.order.same == TRUE){
  ave_expression_matrix_raw <- (rep1_shared_gene_matrix　+ rep2_shared_gene_matrix)/2
}else{
  print("Value Err!")
}

#Exclude na containing gene.
medaka_2y_RNAseq_ave <- ave_expression_matrix_raw[complete.cases(ave_expression_matrix_raw),]
medaka_2y_RNAseq_s1 <- rep1_matrix[complete.cases(rep1_matrix),]
medaka_2y_RNAseq_s2 <- rep2_matrix[complete.cases(rep2_matrix),]

# Note: Average expression matrix losts NA information
print(dim(medaka_2y_RNAseq_s1))
print(dim(medaka_2y_RNAseq_s2))
print(dim(medaka_2y_RNAseq_ave))

# Normalize data
medaka_2y_RNAseq_ave_norm <- scale(t(medaka_2y_RNAseq_ave)) %>% t()
print(medaka_2y_RNAseq_ave_norm[1:5,1:5])
```

### Loading Signal Importance.

Load signal importance generated in `InverseSignalImportance_share.ipynb`. See method for detail estimation process. Convert twice per month data into once per month to align with the RNA-seq data.

```{r LoadSignalImportance_room}

path　= "../results/SignalImportance_average.csv"


signal_importance_ <- read.csv(path, row.names = 1)

#Align the sample date
sample_point <- format(as.Date(signal_importance_$Time_raw, format = "%Y-%m-%d %H:%M:%S"), "%y-%m/%d")
signal_importance <- signal_importance_[is.element(sample_point, colnames(medaka_2y_RNAseq_ave)),]
head(signal_importance)
```

### Comparison of RNA-seq and Signal importance.

Compare the gene you are interested in and signal importance.

```{r ComparisonRNAseqSignalImportance}

comp_plot <- function(symbol, factor){
  factor_color <- c(SR = "red", WT = "blue", DL = "orange", GSI = "#008000")
  geneid = filter(symbol_to_geneid, Symbol == symbol)$gene_id
  if(length(geneid) == 0){
    geneid = symbol
  }
  comp_df <- data.frame(sample_point = c(1:24),
                        gene = scale(as.numeric(medaka_2y_RNAseq_ave_norm[geneid,])),
                        sig_imp = scale(as.numeric(signal_importance[[factor]])))
  g <- 
    ggplot(comp_df) +
    geom_line(aes(x = sample_point, y = gene, color = "Expression"), 
              alpha = 0.6, size = 1) + 
    geom_point(aes(x = sample_point, y = gene, color = "Expression"), 
               alpha = 0.6, size = 1) + 
    geom_line(aes(x = sample_point, y = sig_imp, color = "Signal Importance"), 
              alpha = 0.6, size = 1) +
    geom_point(aes(x = sample_point, y = sig_imp, color = "Signal Importance"), 
               alpha = 0.6, size = 1) +
    labs(x = "Sample Date", y = "Value", title = "") +
    scale_color_manual(
        name = "",
        values = c("Expression" = "black", "Signal Importance" = factor_color[[factor]]),  
        labels = c("Expression" = symbol, "Signal Importance" = "W_WT") 
    ) +
    scale_x_continuous(
      breaks = seq(1, 26, by = 2),  # ラベルを設定する位置
      labels = c("10", "12", "02", "04", "06", "08", "10", "12", "02", "04", "06", "08", "10")
    ) + 
    ylab("Normalize") + xlab("Time (month)") + 
    theme(
      legend.direction = "vertical",
      legend.position = "inside", legend.position.inside =  c(.01, .95),
      legend.justification = "left",
      legend.background = element_blank(), 
      legend.text = element_text(size = 12)
    )
  return(g)
}

#g1 <- comp_plot("drg1", "SR")
#g2 <- comp_plot("LOC101160627", "WT")
#g3 <- comp_plot("LOC110013827", "GSI")
#gblock <- gridExtra::grid.arrange(g1,g2,g3, nrow = 1,ncol=3)

g1 <- comp_plot("LOC101160627", "WT")
g2 <- comp_plot("LOC101162833", "WT")
g3 <- comp_plot("BGI_novel_G000005", "WT")
g4 <- comp_plot("BGI_novel_G000075", "WT")

gblock <- gridExtra::grid.arrange(g1,g2,g3,g4,nrow = 2,ncol=2)
#ggplot2::ggsave('MedakaPosterFig.pdf',gblock, 
#               width=9, height =18, units = "cm", dpi=200)

```

# Statistical test

### Loading randomized signal importance.

Load surrogated signal importance generated in `RandomizeSignalImportance.ipynb`.

```{r LoadRandomizedWaveform}

data_dir = "../results"
SRimp.randomized.coef_ <- read.csv(file.path(data_dir, paste("randomized_SRimp.csv", sep="")), header=FALSE)
WTimp.randomized.coef_ <- read.csv(file.path(data_dir, paste("randomized_WTimp.csv", sep="")), header=FALSE)
DLimp.randomized.coef_ <- read.csv(file.path(data_dir, paste("randomized_DLimp.csv", sep="")), header=FALSE)
GSIimp.randomized.coef_ <- read.csv(file.path(data_dir, paste("randomized_GSIimp.csv", sep="")), header=FALSE)


#Align the sample date
sample_point <- format(as.Date(signal_importance_$Time_raw, format = "%Y-%m-%d %H:%M:%S"), "%y-%m/%d")

SRimp.randomized.coef <- SRimp.randomized.coef_[,is.element(sample_point, colnames(medaka_2y_RNAseq_ave))]
WTimp.randomized.coef <- WTimp.randomized.coef_[,is.element(sample_point, colnames(medaka_2y_RNAseq_ave))]
DLimp.randomized.coef <- DLimp.randomized.coef_[,is.element(sample_point, colnames(medaka_2y_RNAseq_ave))]
GSIimp.randomized.coef <- GSIimp.randomized.coef_[,is.element(sample_point, colnames(medaka_2y_RNAseq_ave))]



#Comparison with gene expression SuppFig5 e
df <- data.frame(
  X = c(1:24), 
  factor = as.numeric(medaka_2y_RNAseq_ave_norm['101160627',]),
  Xcoef = scale(as.numeric(signal_importance[["WT"]])),
  Xcoef1 = scale(as.numeric(WTimp.randomized.coef[75290,])),
  Xcoef2 = scale(as.numeric(WTimp.randomized.coef[22661,])),
  Xcoef3 = scale(as.numeric(WTimp.randomized.coef[18446,]))
) %>% 
  pivot_longer(-X, names_to = "key", values_to = "value")

color_map <- c("factor" = "black", "Xcoef" = "blue", "Xcoef1" = "coral", 
               "Xcoef2" = "#008080", "Xcoef3" = "goldenrod")
label_map <- c("factor" = '101160627', "Xcoef" = "W_WT", "Xcoef1" = 'surrogate #75289', 
               "Xcoef2" = 'surrogate #22660', "Xcoef3" = 'surrogate #18445')

ggplot(df, aes(x = X, y = value, color = key)) +  
  geom_line() +
  scale_color_manual(name = "", values = color_map, labels = label_map) +
  scale_x_continuous(breaks = seq(1, 26, by = 2), 
                     labels =c("10","12","02","04","06","08","10","12","02","04","06","08","10")) +
  theme_classic(base_size = 12) + 
  labs(y = "FPKM · Weight", x = "Time (month)") +
  theme(legend.position = "top", legend.text = element_text(size = 12))


```

### Randomization test.

In statistical test, we only focus on the highly correlate genes (r \> 0.6).

#### Surrogate method explaining

```{r nullmodeltest}

nullmodel.ccf <- cor(t(WTimp.randomized.coef), as.numeric(medaka_2y_RNAseq_ave_norm['101160627', ]))
realdata.ccf <- cor(as.numeric(signal_importance[["WT"]]),as.numeric(medaka_2y_RNAseq_ave_norm['101160627',]))
g_mat <-
  ggplot(data.frame(x = nullmodel.ccf), aes(x = x)) +
  geom_histogram(aes(y = ..density..), fill = "gray", color = "grey", alpha = 0.8) +
  geom_density(color = "black", linewidth = 1) +
  geom_vline(xintercept = 0.8, color = "red", linetype = "dashed", linewidth = 1) +
  theme_classic(base_size = 12) +
  ggtitle('101160627') +
  labs(x = "Correlation coefficient", y = "Density")

dens <- density(nullmodel.ccf)

data_point = which(dens$x < realdata.ccf)
bin_width = diff(dens$x)[1]
if(realdata.ccf > 0){
  p_val <- 1- sum(dens$y[data_point])*bin_width
}else{
  p_val <- sum(dens$y[data_point])*bin_width
}
print(p_val)
g_mat
```

#### Quantify correlation coefficients for signal importance and each gene.

```{r DefinePermutationTestFinc}

## Filtering sample 

SRimp_ccr <- apply(medaka_2y_RNAseq_ave_norm, 1, y = signal_importance[['SR']], FUN=cor) %>% sort(decreasing = T)　
WTimp_ccr <- apply(medaka_2y_RNAseq_ave_norm, 1, y = signal_importance[['WT']], FUN=cor) %>% sort(decreasing = T)
DLimp_ccr <- apply(medaka_2y_RNAseq_ave_norm, 1, y = signal_importance[['DL']], FUN=cor) %>% sort(decreasing = T)
GSIimp_ccr <- apply(medaka_2y_RNAseq_ave_norm, 1, y = signal_importance[['GSI']], FUN=cor) %>% sort(decreasing = T)
#SRimp.vs.sexgene = sapply(rownames(sex_hormone_genes), nullmodeldf=SRimp.randomized.coef, realmodeldf=signal_importance[['SR']], FUN = surrogate_test) %>% t()

SRtmp = SRimp_ccr %>% as.data.frame()
colnames(SRtmp) = "cor"
WTtmp = WTimp_ccr %>% as.data.frame()
colnames(WTtmp) = "cor"
DLtmp = DLimp_ccr %>% as.data.frame()
colnames(DLtmp) = "cor"
GSItmp = GSIimp_ccr %>% as.data.frame()
colnames(GSItmp) = "cor"

histSR <- SRtmp %>%   ggplot(aes(x=cor)) + geom_histogram(colour="black", fill="grey", linewidth=0.1) + xlim(-1,1) + ggtitle(bquote(w[.("SR")])) + xlab("Correlation coefficient")

histWT <-WTtmp %>% ggplot(aes(x=cor)) + geom_histogram(colour="black", fill="grey", linewidth=0.1) + xlim(-1,1) + ggtitle(bquote(w[.("WT")])) + xlab("Correlation coefficient")

histDL <-DLtmp %>% ggplot(aes(x=cor)) + geom_histogram(colour="black", fill="grey", linewidth=0.1) + xlim(-1,1) + ggtitle(bquote(w[.("DL")])) + xlab("Correlation coefficient")

histGSI <-GSItmp %>% ggplot(aes(x=cor)) + geom_histogram(colour="black", fill="grey", linewidth=0.1) + xlim(-1,1) + ggtitle(bquote(w[.("GSI")])) + xlab("Correlation coefficient")

histgblock <- gridExtra::grid.arrange(histSR,histWT,histGSI, nrow=3)
```

```{r correlation_plot}

comp_plot <- function(geneid, factor){
  
  factor_color <- c(SR = "red", WT = "blue", DL = "orange", GSI = "#008000")
  symbol = filter(symbol_to_geneid, gene_id==geneid)$Symbol
  if(is.na(symbol)){
    symbol = geneid
  }
  comp_df <- data.frame(sample_point = c(1:24),
                        gene = scale(as.numeric(medaka_2y_RNAseq_ave_norm[geneid,])),
                        sig_imp = scale(as.numeric(signal_importance[[factor]])))
  g <- 
    ggplot(comp_df) +
    geom_line(aes(x = sample_point, y = gene, color = "Expression"), 
              alpha = 0.6, size = 0.5) + 
    geom_point(aes(x = sample_point, y = gene, color = "Expression"), 
               alpha = 0.6, size = 0.5) + 
    geom_line(aes(x = sample_point, y = sig_imp, color = "Signal Importance"), 
              alpha = 0.6, size = 0.5) +
    geom_point(aes(x = sample_point, y = sig_imp, color = "Signal Importance"), 
               alpha = 0.6, size = 0.5) +
    labs(x = "Sample Date", y = "Value", title = "") +
    scale_color_manual(
        name = "",
        values = c("Expression" = "black", "Signal Importance" = factor_color[[factor]]),  
        breaks="Expression",
        labels=symbol
    ) +
    scale_x_continuous(
      breaks = seq(1, 26, by = 4),  # ラベルを設定する位置
      labels = c("10", "02", "06", "10", "02", "06", "10")
    ) + 
    ylab("Normalize") + xlab("Time (month)") + 
    theme(
      legend.direction = "vertical",
      legend.position = "inside", legend.position.inside =  c(.01, .95),
      legend.justification = "left",
      legend.background = element_blank(), 
      legend.text = element_text(size = 6)
    )
  return(g)
}

gSR <- gridExtra::grid.arrange(
  grobs = lapply(names(SRimp_ccr[1:3]), comp_plot, "SR"),
  ncol = 3
)
interest <- c("BGI_novel_G000005", "BGI_novel_G000075", "101159287")
gWT <- gridExtra::grid.arrange(
  grobs = lapply(interest, comp_plot, "WT"),
  ncol = 3
)
gDL <- gridExtra::grid.arrange(
  grobs = lapply(names(DLimp_ccr[1:3]), comp_plot, "DL"),
  ncol = 3
)
gGSI <- gridExtra::grid.arrange(
  grobs = lapply(names(GSIimp_ccr[1:3]), comp_plot, "GSI"),
  ncol = 3
)

compgblock <- gridExtra::grid.arrange(gSR,gWT,gGSI, nrow=3)

library(grid)   
Fig3abplot <- gridExtra::grid.arrange(compgblock, histgblock, ncol=2, widths=c(3,0.8))

#ggplot2::ggsave('FiguresV2/plot3ab.pdf',Fig3abplot, 
#               width=7.13, height =3.47, units = "in")


```

#### Calculate p-value based on surrogate test

```{r PermutationTest}

# Filtering interest gene name sets
SRimp_highcc_geneids <- names(SRimp_ccr[SRimp_ccr > 0.6])
WTimp_highcc_geneids <- names(WTimp_ccr[WTimp_ccr > 0.6])
DLimp_highcc_geneids <- names(DLimp_ccr[DLimp_ccr > 0.6])
GSIimp_highcc_geneids <- names(GSIimp_ccr[GSIimp_ccr > 0.6])


library(matrixStats)
library(future.apply)
library(progressr)   # ← 進捗表示

surrogate_test <- function(genename,
                           nullmodel_mat,   # 行 = サロゲート (N=1e4), 列 = 時点
                           realmodel_vec) { # 長さ T
  ## 1. 実データ相関
  g     <- as.numeric(medaka_2y_RNAseq_ave_norm[genename, ])
  realr <- cor(realmodel_vec, g)
  ## 2. サロゲートとの相関を一括計算
  null_r <- cor(t(nullmodel_mat), g)

  #公式（Davison & Hinkley, 1997）：p = (k+1)/(N+1)
  # calculate p-value from the emperical distribution.
  N    <- length(null_r)
  k_hi <- sum(null_r >=  realr)
  k_lo <- sum(null_r <=  realr)
  pval <- if (realr >= 0) (k_hi + 1) / (N + 1)
           else            (k_lo + 1) / (N + 1)
  return(c(cor = realr, p_val = pval))
}


# fast mode sapply
options(mc.cores = parallel::detectCores())

SRimp_hc_statres <- pbapply::pbsapply(
  SRimp_highcc_geneids,
  surrogate_test,
  nullmodel_mat = SRimp.randomized.coef,
  realmodel_vec = as.numeric(signal_importance[['SR']])
)  %>% t() %>% as.data.frame()
colnames(SRimp_hc_statres) <- c("cor", "raw_pval")

WTimp_hc_statres <- pbapply::pbsapply(
  WTimp_highcc_geneids,
  #rownames(medaka_2y_RNAseq_ave),
  surrogate_test,
  nullmodel_mat = WTimp.randomized.coef,
  realmodel_vec = as.numeric(signal_importance[['WT']])
) %>% t() %>% as.data.frame()
colnames(WTimp_hc_statres) <- c("cor", "raw_pval")

DLimp_hc_statres <- pbapply::pbsapply(
  DLimp_highcc_geneids,
  surrogate_test,
  nullmodel_mat = DLimp.randomized.coef,
  realmodel_vec = as.numeric(signal_importance[['DL']])
) %>% t() %>% as.data.frame()
colnames(DLimp_hc_statres) <- c("cor", "raw_pval")

GSIimp_hc_statres <- pbapply::pbsapply(
  GSIimp_highcc_geneids,
  surrogate_test,
  nullmodel_mat = GSIimp.randomized.coef,
  realmodel_vec = as.numeric(signal_importance[['GSI']])
) %>% t() %>% as.data.frame()
colnames(GSIimp_hc_statres) <- c("cor", "raw_pval")

```

#### Extract significantly correlate genes

```{r signif_gene}
# Multiple correction.
SRimp_hc_statres_corrected <- SRimp_hc_statres %>% 
  mutate(bonfp = p.adjust(raw_pval, method = "bonferroni"))

WTimp_hc_statres_corrected <- WTimp_hc_statres %>% 
  mutate(bonfp = p.adjust(raw_pval, method = "bonferroni"))

DLimp_hc_statres_corrected <- DLimp_hc_statres %>% 
  mutate(bonfp = p.adjust(raw_pval, method = "bonferroni"))

GSIimp_hc_statres_corrected <-GSIimp_hc_statres %>%
  mutate(bonfp = p.adjust(raw_pval, method = "bonferroni"))

# Filter with p < 0.05 (bonferroni)
SRimp_hc_statres_corrected %>% filter(bonfp < 0.05) %>% print() 
WTimp_hc_statres_corrected %>% filter(bonfp < 0.05) %>% print() # 4 genes
DLimp_hc_statres_corrected %>% filter(bonfp < 0.05) %>% print() # 0.05: 5 genes, 0.01: 2 genes
GSIimp_hc_statres_corrected %>% filter(bonfp < 0.05) %>% print() # 0 genes

# Extract significant correlated genes
WTimp_signif_geneids <- WTimp_hc_statres_corrected %>% filter(bonfp < 0.05) %>% rownames() %>% sort()
SRimp_signif_geneids <- SRimp_hc_statres_corrected %>% filter(bonfp < 0.05) %>% rownames() 
DLimp_signif_geneids <- DLimp_hc_statres_corrected %>% filter(bonfp < 0.05) %>% rownames() 
# show detail
gene_profiles %>% filter(is.element(gene_id, WTimp_signif_geneids)) %>% 
  mutate(name = ifelse(is.na(Symbol), gene_id, Symbol)) %>% 
  dplyr::select(name) %>% gt()
```

```{r create_suptable_1}

if(!requireNamespace("writexl", quietly = TRUE)){install.packages("writexl")}
library(writexl)

SR_save_format <- SRimp_hc_statres_corrected %>% rownames_to_column() %>% 
  set_names(c("gene_id", "cor with W_SR", "p-val", "p_adj (bonferroni)")) %>% 
  left_join(gene_profiles)
WT_save_format <- WTimp_hc_statres_corrected %>% rownames_to_column() %>% 
  set_names(c("gene_id", "cor with W_WT", "p-val", "p_adj (bonferroni)")) %>% 
  left_join(gene_profiles)
DL_save_format <- DLimp_hc_statres_corrected %>% rownames_to_column() %>% 
  set_names(c("gene_id", "cor with W_DL", "p-val", "p_adj (bonferroni)")) %>% 
  left_join(gene_profiles)
GSI_save_format <- GSIimp_hc_statres_corrected %>% rownames_to_column() %>% 
  set_names(c("gene_id", "cor with W_GSI", "p-val", "p_adj (bonferroni)")) %>% 
  left_join(gene_profiles)


output_path <- "SupplementTable1.xlsx"
# Export supplementary table for each signal importance to Excel.
# Skipped if the file already exists to prevent accidental overwriting.
if (!file.exists(output_path)) {
  write_xlsx(
    list(
      SRimp  = SR_save_format,
      WTimp  = WT_save_format,
      DLimp  = DL_save_format,
      GSIimp = GSI_save_format
    ),
    path = output_path
  )
  message("Saved: ", output_path)
} else {
  message("Already exists, skipped: ", output_path)
}
```

# Gene Ontology analysis of highly correlate genes

0.  Create medaka GO database. See Nakayama et al., 2023 for GO and KO annotation method.
1.  GO analysis for correlate genes
2.  KO analysis for correlate genes

### Create database

```{r create_database}
library(AnnotationForge)
fSym <- gene_profiles %>% dplyr::select(gene_id, Symbol) %>% 
  mutate(Symbol = ifelse(is.na(Symbol), gene_id, Symbol))
colnames(fSym) <- c("GID","SYMBOL")

fGO_CC <- gene_profiles %>% 
  dplyr::select(gene_id, `Cellular.Component`) %>%
  filter(!is.na(`Cellular.Component`)) %>%  # GOtermがNAでない行を選択
  separate_rows(`Cellular.Component`, sep = ";") %>%  # GOtermが複数ある場合に分割
  mutate(`Cellular.Component` = gsub("//.*", "", `Cellular.Component`))  %>% 
  filter(`Cellular.Component`　!= "NA")
colnames(fGO_CC) <- c("gene_id", "GO")

fGO_MC <- gene_profiles %>% 
  dplyr::select(gene_id, `Molecular.Function`) %>%
  filter(!is.na(`Molecular.Function`)) %>%  # GOtermがNAでない行を選択
  separate_rows(`Molecular.Function`, sep = ";") %>%  # GOtermが複数ある場合に分割
  mutate(`Molecular.Function` = gsub("//.*", "", `Molecular.Function`))  %>% 
  filter(`Molecular.Function`　!= "NA")
colnames(fGO_MC) <- c("gene_id", "GO")

fGO_BP <- gene_profiles %>% 
  dplyr::select(gene_id, `Biological.Process`) %>%
  filter(!is.na(`Biological.Process`)) %>%  # GOtermがNAでない行を選択
  separate_rows(`Biological.Process`, sep = ";") %>%  # GOtermが複数ある場合に分割
  mutate(`Biological.Process` = gsub("//.*", "", `Biological.Process`))  %>% 
  filter(`Biological.Process`　!= "NA") 
colnames(fGO_BP) <- c("gene_id", "GO")

fGO <- rbind(fGO_CC, fGO_MC, fGO_BP) %>% mutate(EVIDENCE = "IEA") #Evidence code, set uniformly IEA as default.
colnames(fGO) <- c("GID","GO", "EVIDENCE")

fKO <- gene_profiles %>% 
  dplyr::select(gene_id, Symbol, `Kegg.Orthology`) %>%
  filter(!is.na(`Kegg.Orthology`)) %>%  # GOtermがNAでない行を選択
  separate_rows(`Kegg.Orthology`, sep = ";") %>%  # GOtermが複数ある場合に分割
  mutate(`Kegg.Orthology` = gsub("//.*", "", `Kegg.Orthology`))  %>% 
  filter(`Kegg.Orthology`　!= "NA") 
colnames(fKO) <- c("GID","Symbol", "KO")

if(!file.exists("./org.Olatipes.eg.db")){
  makeOrgPackage(gene_info=fSym, go=fGO,
                 version="0.2",
                 maintainer="Medaka-Ninja <so@someplace.org>",
                 author="Medaka-Samurai <so@someplace.org>",
                 outputDir = ".",
                 tax_id="8090",
                 genus="Oryzias",
                 species="latipes",
                 goTable="go")
  ## then you can call install.packages based on the return value
}
```

### GO analysis.

```{r GO_entichment}
if(!requireNamespace("./org.Olatipes.eg.db", quietly = TRUE)){install.packages("./org.Olatipes.eg.db", repos=NULL, type="source")}
library(org.Olatipes.eg.db)

enrich_plot <- function(list, ont, col, title_){
  ego_result <- clusterProfiler::enrichGO(
                           gene=list,
                           OrgDb=org.Olatipes.eg.db,
                           keyType = "GID",
                           ont=ont, 　　　　　　　　　　　　#"BP","CC","MF","ALL"から選択
                           pAdjustMethod="fdr",
                           pvalueCutoff=0.1, # 0.1 returns too much result.
                           qvalueCutoff=0.1, 
                           readable=TRUE)
  #Tests must pass i)
  #pvalueCutoff on unadjusted pvalues, ii) pvalueCutoff on adjusted pvalues
  #and iii) qvalueCutoff on qvalues to be reported.
  #https://bioconductor.org/packages/devel/bioc/manuals/clusterProfiler/man/clusterProfiler.pdf
  
  if(nrow(as.data.frame(ego_result)) < 1){
    print("no enruchment")
      return(NA)
  }
  ego_result@result = ego_result@result %>% mutate(log10pvalue = log10(pvalue))
  min_x = min(ego_result@result$log10pvalue)
  print(ego_result@result)
  
  fs = 12
  # fs = 5 when export to illustorator. 
  g <- graphics::barplot(ego_result, x="log10pvalue", showCategory=dim(ego_result)[1], 
                         font.size = fs, label_format = 50) + 
    xlim(0,min_x) +
    labs(x = "qscore", y = " ") +
    scale_fill_gradientn(colors = c(col, col)) + 
    theme(legend.position = "none", 
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank()) + 
    ggtitle(title_) + 
    theme(plot.title = element_text(size = 6))

  return(g)
}


BPSR <- enrich_plot(SRimp_highcc_geneids, "BP", "#EC2526", "W_sr high correlate: BP") 
BPWT <-enrich_plot(WTimp_highcc_geneids, "BP", "#3B54A4", "W_wt high correlate: BP")
BPDL <-enrich_plot(DLimp_highcc_geneids, "BP", "orange", "W_DL high correlate: BP") #NA
BPGSI <-enrich_plot(GSIimp_highcc_geneids, "BP", "#0F8141", "W_gsi high correlate: BP") #NA

MFSR <- enrich_plot(SRimp_highcc_geneids, "MF", "#EC2526", "W_sr high correlate: MF") 
MFWT <-enrich_plot(WTimp_highcc_geneids, "MF", "#3B54A4", "W_wt high correlate: MF")
MFDL <-enrich_plot(DLimp_highcc_geneids, "MF", "orange", "W_DL high correlate: MF") #NA
MFGSI <-enrich_plot(GSIimp_highcc_geneids, "MF", "#0F8141", "W_gsi high correlate: MF") #NA


MFWTsignif <- enrich_plot(WTimp_signif_geneids, "MF", "darkblue", "W_wt signif. high correlate: MF")
BPWTsignif <- enrich_plot(WTimp_signif_geneids, "BP", "darkblue", "W_wt signif. high correlate: BP")


highccGO <- gridExtra::grid.arrange(MFSR, MFWT, BPWT,heights = c(1.1, 1, 1.8))
sigccGO <-  gridExtra::grid.arrange(MFWTsignif, BPWTsignif,  heights = c(1.0, 0.9))

GOresult <- gridExtra::grid.arrange(highccGO, sigccGO, ncol=2) 

#copy and paste to the console.

{
  plot(GOresult)
  grid.text("a", x = unit(0.01, "npc"), y = unit(1, "npc"),
            just = c("left", "top"), gp = gpar(fontsize = 7, fontface = "bold"))
  grid.text("b", x = unit(0.01, "npc"), y = unit(0.8, "npc"),
          just = c("left", "top"), gp = gpar(fontsize = 7, fontface = "bold"))
  grid.text("c", x = unit(0.01, "npc"), y = unit(0.63, "npc"),
          just = c("left", "top"), gp = gpar(fontsize = 7, fontface = "bold"))
  grid.text("d", x = unit(0.01, "npc"), y = unit(0.3, "npc"), 
            just = c("left", "top"), gp = gpar(fontsize = 7, fontface = "bold"))
  grid.text("e", x = unit(0.01, "npc"), y = unit(0.1, "npc"), 
            just = c("left", "top"), gp = gpar(fontsize = 7, fontface = "bold"))
  grid.text("f", x = unit(0.55, "npc"), y = unit(1, "npc"), 
            just = c("left", "top"), gp = gpar(fontsize = 7, fontface = "bold"))
  grid.text("g", x = unit(0.55, "npc"), y = unit(0.48, "npc"), 
            just = c("left", "top"), gp = gpar(fontsize = 7, fontface = "bold"))
}


# Since the figure is large, we recommend viewing it in the plot pane
# before exporting. Uncomment the lines below to save as PDF.
# ggplot2::ggsave(
#   filename = "FiguresV2/Fig3raw.pdf",
#   plot     = fig3,
#   width    = 7.13,
#   height   = 6.7,
#   units    = "in"
# )

plot(GOresult)
```

### KO analysis.

Although single gene may be associated with multiple K numbers, since the background data shares the same characteristics, this does not pose a problem.

```{r KO_enrichment}
 KO_enrich_plot <- function(target_Knumbers, col){
    KOe_result <- clusterProfiler::enrichKEGG(
      target_Knumbers,
      organism = "ko", 
      # "you can use enrichKEGG with K number by specifying organism="ko". 
      # https://github.com/YuLab-SMU/clusterProfiler/issues/99
      keyType = "kegg",
      pvalueCutoff = 0.1,
      qvalueCutoff = 0.1,
      pAdjustMethod = "fdr",
      minGSSize = 10,
      maxGSSize = 500,
      universe = fKO$KO,
      use_internal_data = FALSE
    )
    
  KOe_result@result = KOe_result@result %>% mutate(log10pvalue = log10(pvalue))
  min_x = min(KOe_result@result$log10pvalue)
  print(paste("enrich:",dim(KOe_result)[1]))
  
  g <- graphics::barplot(KOe_result, x="log10pvalue", showCategory=dim(KOe_result)[1], 
                         font.size = 12, label_format = 50) + 
    xlim(0,min_x) +
    labs(x = "-log10(pvalue)", y = "") +
    scale_fill_gradientn(colors = c(col, col)) + 
    theme(legend.position = "none", 
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank())

  return(g)
}


SR_knumbers <- filter(fKO, is.element(GID, SRimp_highcc_geneids))$KO
WT_knumbers <- filter(fKO, is.element(GID, WTimp_highcc_geneids))$KO
DL_knumbers <- filter(fKO, is.element(GID, DLimp_highcc_geneids))$KO
GSI_knumbers <- filter(fKO, is.element(GID, GSIimp_highcc_geneids))$KO
WTsignif_knumbers <- filter(fKO, is.element(GID, WTimp_signif_geneids))$KO

KeggSR <-  KO_enrich_plot(SR_knumbers, "#EC2526") #NA
KeggWT <- KO_enrich_plot(WT_knumbers, "#3B54A4") #NA
KeggDL <- KO_enrich_plot(DL_knumbers, "orange") #NA
KeggGSI <- KO_enrich_plot(GSI_knumbers, "#0F8141") #NA
#g1 <- gridExtra::grid.arrange(KeggSR,KeggWT,KeggGSI, nrow=3,ncol=1, heights=c(1.5,0.8,0.6))

KO_enrich_plot(WTsignif_knumbers, "darkblue")
#KO_enrich_plot(SRsignif_knumbers, "#EC2526")
#KO_enrich_plot(DLsignif_knumbers, "orange")

KOe_result <- clusterProfiler::enrichKEGG(
    WTsignif_knumbers,
    organism = "ko", 
    keyType = "kegg",
    pvalueCutoff =0.1,
    qvalueCutoff = 0.1,
    pAdjustMethod = "fdr",
    universe = fKO$KO,
    minGSSize = 10,
    maxGSSize = 500,
    use_internal_data = FALSE
)
namekodf <- fKO %>% filter(is.element(GID, WTimp_signif_geneids)) %>% 
  mutate(Symbol = ifelse(is.na(Symbol), GID, Symbol))

symbols = sapply(KOe_result@result$geneID,function(ko){
  v1 = namekodf %>% filter(is.element(KO, strsplit(ko, "/")[[1]]))
  v2 = unique(v1$Symbol)
  return(paste(toString(v2), sep=","))
}) %>% unlist()



KOe_result@result <- cbind(KOe_result@result, symbols)
KOe_result@result = KOe_result@result %>% mutate(log10pvalue = log10(pvalue))
min_x = min(KOe_result@result$log10pvalue)
KeggSignifWT <- 
   graphics::barplot(KOe_result, x="log10pvalue", showCategory=dim(KOe_result)[1], 
                         font.size = 5, label_format = 20) + 
    xlim(0,min_x) +
    labs(x = "-log10(pvalue)", y = "") +
    theme(legend.position = "none", 
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank()) +
  geom_text(aes(label=symbols, x=-0.04), size = 1, color = "white",
            show.legend = FALSE,
            position = position_dodge(0),
            hjust = 0) + 
  scale_fill_gradientn(colors = c("darkblue", "darkblue")) + 
  ggtitle("W_wt significantly correlate: KO") +
   theme(plot.title = element_text(size = 6))

plot(KeggSignifWT)
```

```{r create_figure2}
highccGOMF <- gridExtra::grid.arrange(MFSR, MFWT, ncol=1, heights=c(1.8,2.8))
fig3cplot <- gridExtra::grid.arrange(highccGOMF, KeggSignifWT, ncol=2) 
fig3 <- gridExtra::grid.arrange(Fig3abplot, fig3cplot,  nrow=2, heights=c(1,1))


{
  #ggplot transaction
  plot(fig3)
  grid.text("a", x = unit(0.01, "npc"), y = unit(1, "npc"),
            just = c("left", "top"), gp = gpar(fontsize = 7, fontface = "bold"))
  grid.text("b", x = unit(0.79, "npc"), y = unit(1, "npc"),
            just = c("left", "top"), gp = gpar(fontsize = 7, fontface = "bold"))
  grid.text("c", x = unit(0.01, "npc"), y = unit(0.5, "npc"),
            just = c("left", "top"), gp = gpar(fontsize = 7, fontface = "bold"))
}

#ggplot2::ggsave('FiguresV2/Fig3tmp.pdf',fig3, 
#               width=7.13, height =6.7, units = "in")
```

# Comparing expression dynamics in room and outdoor conditions.

### Loading RNA-seq data: GSE234565

Transcriptome data collected under stable indoor conditions (constant environment, no seasonal variation).\
GEO accession: <https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE234565>

```{r LoadRoomRNAseq}
library(readxl)　

# Automatically download.
# If this does not work, download excel file manually from below link.
#url = "https://ftp.ncbi.nlm.nih.gov/geo/series/GSE234nnn/GSE234565/suppl#/GSE234565%5FFPKM%5Fmedaka%5Fconstant%5FRNA%2Dseq%2Exlsx"
#temp_file <- tempfile(fileext = ".xlsx")
#curl::curl_download(url, temp_file)

temp_file <- "GSE234565_FPKM_medaka_constant_RNA-seq.xlsx"
print(excel_sheets(temp_file))  # AllSamples.GeneExpression.FPKM_
medaka_room_RNAseq_raw <- read_excel(temp_file, sheet = 1)
medaka_room_RNAseq <- medaka_room_RNAseq_raw %>% 
  dplyr::select(-"SymbolID", -"transcript_id(s)", -"Cellular Component", -"Molecular Function", -"Biological Process", -"Kegg Orthology", -"Nr Description") %>% 
  column_to_rownames("gene_id")


head(medaka_room_RNAseq)
```

### Data formatting

Samples are named according to the number of days elapsed since transfer to indoor conditions. The exact post-transfer day for each sample is listed in its metadata　(<https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSM7471033>).\
Collection month is indicated in Fig. S1E of Nakayama et al., 2023 (manually stored in the variable `sample_month`) .

```{r DataFormattingRoomRNAseq}

# data attribute.
days_after_transition <- c(0, 0, 0, 7, 7, 7, 30, 30, 30, 60, 60, 90, 90, 90, 120, 120, 120, 150, 150, 150)
sample_labels <- c("sample1", "sample2", "sample3", "sample1", "sample2", "sample3", "sample1", "sample2", "sample3", "sample1", "sample2","sample1", "sample2",
                   "sample3", "sample1", "sample2", "sample3", "sample1", "sample2", "sample3")
sample_month <- c(3, 3, 3, 3.2, 3.2, 3.2, 4, 4, 4, 5, 5, 6, 6, 6, 7, 7, 7, 8, 8, 8)

colnames(medaka_room_RNAseq) <- sample_labels
sample1 <- medaka_room_RNAseq[,colnames(medaka_room_RNAseq) == "sample1"] 
sample2 <- medaka_room_RNAseq[,colnames(medaka_room_RNAseq) == "sample2"] 
sample3_ <- medaka_room_RNAseq[,colnames(medaka_room_RNAseq) == "sample3"] 

colnames(sample1) <- days_after_transition[which(sample_labels == "sample1")]
colnames(sample2) <- days_after_transition[which(sample_labels == "sample2")]
colnames(sample3_) <- days_after_transition[which(sample_labels == "sample3")] 

# “60 days after transition”  in Sample 3 which have no data is filled with missing values.
sample3 <- mutate(sample3_, "60"=NA) 
column_order = colnames(sample3) %>% as.numeric() %>% sort() %>% as.character()
sample3 <- sample3[,column_order]

# Taking average among sample
medaka_room_RNAseq_merge <- rbind(rownames_to_column(sample1), rownames_to_column(sample2), rownames_to_column(sample3)) %>% group_by(rowname)
medaka_room_RNAseq_ave <- summarize(medaka_room_RNAseq_merge , "0"=mean(`0`), "7"=mean(`7`), "30"=mean(`30`), "60"=mean(`60`, na.rm=TRUE), "90"=mean(`90`), "120"=mean(`120`), "150"=mean(`150`)) %>% 
  column_to_rownames("rowname") 





# Extract outdoor samples corresponding to the period (Mar~Aug) during which indoor samples were collected.

FIRST_YEAR_INDEX = c(6:11)
#SECOND_YEAR_INDEX = c(18:23)
medaka_outdoor_RNAseq_s1 = medaka_2y_RNAseq_s1[,FIRST_YEAR_INDEX]
medaka_outdoor_RNAseq_s2 = medaka_2y_RNAseq_s2[,FIRST_YEAR_INDEX]
medaka_outdoor_RNAseq_ave = medaka_2y_RNAseq_ave[,FIRST_YEAR_INDEX]
```

```{r testplot}

plot3_sample_v2 <- function(genename){
    
  # Data frame creation
  symb = filter(symbol_to_geneid, gene_id == genename)$Symbol
  if(is.na(symb)){symb = genename}
  g <- data.frame(time = unique(sample_month),
                  meanval = as.numeric(medaka_room_RNAseq_ave[genename,]),
                  sample1 = as.numeric(sample1[genename,]),
                  sample2 = as.numeric(sample2[genename,]),
                  sample3 = as.numeric(sample3[genename,])
  ) %>% ggplot() + 
    geom_line(aes(x=time, y=meanval, color="Room")) +
    geom_point(aes(x=time, y=sample1, color="Room"), shape=20,size=0.8) + 
    geom_point(aes(x=time, y=sample2, color="Room"), shape=20,size=0.8) + 
    geom_point(aes(x=time, y=sample3, color="Room"), shape=20,size=0.8) + 
    ylab("FPKM") + 
    xlab("Time (month)") + 
    ggtitle(symb)
  

  # Adding points for outdoor samples
  
  OutPoint <- data.frame(
    outtime= c(3:8),
    meanval =  as.numeric(medaka_outdoor_RNAseq_ave[genename,]),
    rep1 = as.numeric(medaka_outdoor_RNAseq_s1[genename,]),
    rep2 = as.numeric(medaka_outdoor_RNAseq_s2[genename,]))
  
  g <- g + 
    geom_point(data = OutPoint, aes(x=outtime, y=rep1, color="Outdoor"), shape=20,size=0.8) + 
    geom_point(data = OutPoint, aes(x=outtime, y=rep2, color="Outdoor"), shape=20,size=0.8) + 
    geom_line(data = OutPoint, aes(x=outtime, y=meanval, color="Outdoor")) + 
    scale_color_manual(values = c("Room" = "purple", "Outdoor" = "black")) + 
    scale_linetype_manual(values = c("Room" = "dashed", "Outdoor" = "solid")) +
    labs() + 
    theme(
      base_size=6,
      legend.position = "none", 
        legend.text = element_text(size = 6)) 
  
  
  # Plot the graph
  return(g)
}

room_out_WT_comp <- gridExtra::grid.arrange(
  grobs = lapply(WTimp_signif_geneids, plot3_sample_v2), ncol = 2
)



```

### Comparing the expression profile in room and outdoor.

```{r room_outdoor_comp}

library(exactRankTests)
## 配色や定数をまとめて管理
COLORS <- list(Room = "purple", Outdoor = "black")

## ──────────────────────────────────────────────────────────
##  1. ボックスプロット ＆ Wilcoxon 検定
## ──────────────────────────────────────────────────────────

library(coin)

make_expr_df <- function(gene_id) {
  # sample 1,2,and 3
  room <- c(sample1[gene_id, ], sample2[gene_id, ], sample3[gene_id, ]) |>
          as.numeric() |>
          tibble(FPKM = _, category = "Room")
  
  # sample 1 and 2
  outdoor <- c(medaka_outdoor_RNAseq_s1[gene_id,], medaka_outdoor_RNAseq_s2[gene_id,]) |>
             as.numeric() |>
             tibble(FPKM = _, category = "Outdoor")

  bind_rows(room, outdoor)
}


plot_box_gene <- function(gene_id) {
  df   <- make_expr_df(gene_id)
  symb <- symbol_to_geneid %>% filter(gene_id == !!gene_id) %>% pull(Symbol) %>% 
          {ifelse(length(.) == 0 | is.na(.), gene_id, .)}
  
  p_val <- wilcox.exact(log2(df$FPKM[df$category == "Room"]),
                        log2(df$FPKM[df$category == "Outdoor"]),
                        paired = FALSE)$p.value
  
  ggplot(df, aes(category, log2(FPKM))) +
    geom_boxplot(aes(color=category, linetype=category), linewidth=0.1, fill=NA, outlier.shape = NA) + 
    geom_jitter(size = 0.2, aes(color=category)) + xlab("") + ggtitle(symb) + 
    scale_fill_manual(values = c("Room" = "purple", "Outdoor" = "black")) + 
    scale_color_manual(values = c("Room" = "purple", "Outdoor" = "black")) + 
    scale_linetype_manual(values = c("Room" = "solid", "Outdoor" = "solid")) +
    theme(legend.position = "none") +
    labs(title = symb,
         subtitle = sprintf("Wilcoxon p = %.3g", p_val),
         x = NULL, y = "log2(FPKM)", size=4) +
    theme(legend.position = "none")
}

## 例：上位 5 遺伝子を並べて描画
room_out_WT_wilcox <- gridExtra::grid.arrange(
  grobs = lapply(WTimp_signif_geneids, plot_box_gene),
  ncol = 4
)


## ──────────────────────────────────────────────────────────
##  2. CCF (相互相関) プロット
## ──────────────────────────────────────────────────────────

mt <- unique(sample_month)
mtout <- c(3.0:8.0)
  
plot_ccf_gene <- function(gene_id) {
  symb <- symbol_to_geneid %>% filter(gene_id == !!gene_id) %>% pull(Symbol) %>% 
          {ifelse(length(.) == 0 | is.na(.), gene_id, .)}
  
  room_vec <- approx(mt,  medaka_room_RNAseq_ave[gene_id, ])$y
  out_vec  <- approx(mtout, medaka_outdoor_RNAseq_ave[gene_id, ])$y
  time_vec <- approx(mtout, medaka_outdoor_RNAseq_ave[gene_id, ])$x
  
  
  p_val <- cor.test(scale(room_vec), scale(out_vec))$p.value
  ccf_dat <- ccf(scale(room_vec)[,1], scale(out_vec)[,1], plot = FALSE)
  
  tibble(lag = ccf_dat$lag*diff(time_vec)[1], acf = ccf_dat$acf) %>% 
    ggplot(aes(lag, acf)) +
    geom_col(fill = "grey60") +
    geom_hline(yintercept = 0, color = "black") +
    labs(title = symb,
         subtitle = sprintf("Spearman p = %.3g", p_val),
         x = "Lag (month)", y = "CCF") +
    ylim(-1, 1) +
    theme(base_size = 6,
          legend.text = element_text(size = 6))
}

room_out_WT_ccf <- gridExtra::grid.arrange(
  grobs = lapply(WTimp_signif_geneids, plot_ccf_gene),
  ncol = 2
)

## ──────────────────────────────────────────────────────────
##  3. 発現レベル・動態類似度テスト
## ──────────────────────────────────────────────────────────

similarity_test <- function(gene_id) {
  room_pt <- as.numeric(c(sample1[gene_id, ], sample2[gene_id, ], sample3[gene_id, ]))
  out_pt  <- as.numeric(c(medaka_outdoor_RNAseq_s1[gene_id, ], medaka_outdoor_RNAseq_s2[gene_id, ]))
  room_ts <- approx(mt,    medaka_room_RNAseq_ave[gene_id, ])$y
  out_ts  <- approx(mtout, medaka_outdoor_RNAseq_ave[gene_id, ])$y
  wil_p  <- wilcox.exact(room_pt, out_pt, paired = FALSE)$p.value
  cor_p <- tryCatch(
    {
      cor.test(
        scale(room_ts),
        scale(out_ts),
      )$p.value
    },
    error = function(e) {
      NA_real_   # エラー時は NA を返す
    }
  )
  log2FD <- log2(mean(out_ts) / mean(room_ts))
  c(wil_p = wil_p, cor_p = cor_p, log2FD = log2FD)
}

## 進捗付き実行
applied_gene <- intersect(rownames(medaka_room_RNAseq_ave), rownames(medaka_outdoor_RNAseq_ave))
sim_res <- pbsapply(applied_gene, similarity_test) |> 
           t() %>%  as.data.frame() %>%  rownames_to_column("gene_id")

## 多重検定補正と注釈
sim_res_corrected <- sim_res %>% 
  mutate(wil_p_adj = p.adjust(wil_p, method = "bonferroni"),
         cor_p_adj = p.adjust(cor_p, method = "bonferroni"),
         is_strict = gene_id %in% c(WTimp_signif_geneids)) %>% 
  left_join(symbol_to_geneid, by = c("gene_id" = "gene_id")) %>% 
  mutate(Symbol = coalesce(Symbol, gene_id)) #if symnols is missing, fill with gene_id.

library(ggrepel)          # ラベル配置

sim_res_corrected_tmp = sim_res_corrected %>% mutate(paint = (wil_p_adj < 0.01 & cor_p_adj >= 0.01))
wilcox_ccf_plot <-
  ggplot(sim_res_corrected_tmp, aes(log10(wil_p_adj), log10(cor_p_adj))) +
  geom_point(alpha = 0.3, size=0.8, aes(colour = paint)) +
  geom_hline(yintercept = -2, linetype = "dashed") +
  geom_vline(xintercept = -2, linetype = "dashed") +
  geom_label_repel(data = filter(sim_res_corrected, is_strict), 
              aes(label = Symbol), 
              size=2,
              box.padding = 1,   # Padding around the text
              point.padding = 0, # Padding between text and point
              segment.color = "black",  # Color of the connecting lines
              segment.size = 0.3) + 
  scale_colour_manual(values = c("grey", "skyblue"), guide = "none") +
  labs(x = "Expression-level similarity (log10 adj-p)",
       y = "Dynamic similarity (log10 adj-p)",
       color = NULL) +
  theme_classic(base_size = 6) 

## 散布図
#circannual_genes <- read.csv("pnas2313514120sd07.csv")
#sim_res_corrected <- sim_res_corrected %>% mutate()

## ──────────────────────────────────────────────────────────
##  4. Fisher 検定（期待 vs 観測 2×2）
## ──────────────────────────────────────────────────────────
fisher_change_test <- function(df, strict_set, p_thres = 0.01) {
  change_all  <- df %>% filter(wil_p_adj < p_thres, cor_p_adj >= p_thres)
  change_str  <- change_all %>% filter(gene_id %in% strict_set)
  valid_test_size <- df  %>% filter(gene_id %in% strict_set) %>% nrow()
  
  tbl <- matrix(c(nrow(change_str),
                  valid_test_size - nrow(change_str),
                  nrow(change_all)   - nrow(change_str),
                  nrow(df) - valid_test_size - (nrow(change_all) - nrow(change_str))),
                nrow = 2, byrow = TRUE,
                dimnames = list(c("Signif.", "All"),
                                c("Change", "No-change")))
  list(table = tbl, fisher = fisher.test(tbl))
}

# speaman(alt: greater), wilcox(alt:both) => p=0.05
ft_out <- fisher_change_test(sim_res_corrected, WTimp_signif_geneids)
print(ft_out$fisher)
print(ft_out$table)


ftdf <- ft_out$table %>% as.data.frame() %>% 
  rownames_to_column("Category") %>% pivot_longer(c(Change, `No-change`))
colnames(ftdf) <- c("Category", "Type", "Value")

ft_plot <- ggplot(ftdf, aes(x = Type, y = Value, fill = Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ Category, scales = "free_y") +  # カテゴリごとに個別のグラフ
  labs(title = "",
       x = "", y = "# Gene") +
  scale_fill_manual(values = c("skyblue", "grey")) +
  theme_minimal() + 
  facet_wrap(~ Category, scales = "free_y") +
  theme(legend.position = "bottom") + theme_classic(base_size=6) + 
  theme(
    base_size=6,
    legend.text = element_text(size = 6),
    legend.position = "None"
  ) 

```

\

### create figure4

```{r create_plotfig4}

fig4row1 <- gridExtra::grid.arrange(nullGrob(), room_out_WT_comp, ncol=2, widths=c(1.8,2.8))
fig4row2 <-  gridExtra::grid.arrange(room_out_WT_wilcox, room_out_WT_ccf, ncol=2, 
                                     widths=c(1,1))

fig4row3 <-  gridExtra::grid.arrange(wilcox_ccf_plot, ft_plot, ncol=2, widths=c(2,1))

fig4 <- gridExtra::grid.arrange(fig4row1, fig4row2, fig4row3, nrow=3, heights=c(1,1,1.2)) 
fig4

#copy and paste to the console.

ggplot2::ggsave('FiguresV2/Fig4raw_0708_v2.pdf',fig4, 
               width=7.13, height =6.7, units = "in")

```

#### additional analysis: circaannual genes identified in the Nakayama et al., 2024.

```{r additionnal}

circannual_genes_df <- read.csv("./pnas2313514120sd07.csv") 

# Oscillating genes in the constant (LW) condition.
#nsSOGs <- read.csv("./pnas2313514120sd06.csv")
#circannual_gene_ids <- intersect(nsSOGs$Gene.ID, LW_nonlinears$Gene.ID)
circannual_gene_ids <- circannual_genes_df$Gene.I

sim_res_corrected_circa <- sim_res_corrected %>% 
  mutate(paint = ifelse((wil_p_adj > 0.01 & cor_p_adj <= 0.01), "q4","other"))  %>% 
  mutate(is_circa = (gene_id %in% circannual_gene_ids))

wilcox_ccf_plot_circa <-
  ggplot(sim_res_corrected_circa, aes(log10(wil_p_adj), log10(cor_p_adj))) +
  geom_point(shape=21, aes(alpha = is_circa, size=is_circa, fill = paint, color=is_circa)) +
  geom_hline(yintercept = -2, linetype = "dashed") +
  geom_vline(xintercept = -2, linetype = "dashed") +
  scale_color_manual(values = c("grey", "black")) +
  scale_fill_manual(values = c(q4="pink", "other"="grey")) +
  scale_alpha_manual(values = c(0.4, 0.8)) +
  scale_size_manual(values = c(1, 3)) +
  labs(x = "Expression-level similarity (log10 adj-p)",
       y = "Dynamic similarity (log10 adj-p)",
       color = NULL) +
  theme_classic(base_size = 6) + 
  theme(legend.position = "none")  # ← これを追加


fisher_coherent_test <- function(df, test_set, p_thres = 0.01) {
  coherent  <- df %>% filter(wil_p_adj > p_thres, cor_p_adj <= p_thres)
  valid_test_size <- df  %>% filter(gene_id %in% test_set) %>% nrow()
  coherent_str  <- coherent %>% filter(gene_id %in% test_set)
  
  tbl <- matrix(c(nrow(coherent_str),
                   valid_test_size - nrow(coherent_str),
                  nrow(coherent)   - nrow(coherent_str),
                  nrow(df) - valid_test_size - (nrow(coherent) - nrow(coherent_str))),
                nrow = 2, byrow = TRUE,
                dimnames = list(c("Circa.", "All"),
                                c("Coherent", "No-coherent")))
  list(table = tbl, fisher = fisher.test(tbl))
}

ft_coherent_circa <- fisher_coherent_test(sim_res_corrected, circannual_gene_ids)
print(ft_coherent_circa$fisher)

ft_coherent_circa_df <- ft_coherent_circa$table %>% as.data.frame() %>% 
  rownames_to_column("Category") %>% 
  pivot_longer(c(Coherent, `No-coherent`)) %>% 
  setNames(c("Category", "Type", "Value"))

#ft_coherent_circa_gg <-
#  ggplot(ft_coherent_circa_df, aes(x = Type, y = Value, fill = Type)) +
#  geom_bar(stat = "identity", position = "dodge") +
#  facet_wrap(~ Category, scales = "free_y") +  # カテゴリごとに個別のグラフ
#  labs(title = "", x = "", y = "# Gene") +
#  scale_fill_manual(values = c("grey", "grey")) +
#  theme_minimal() + 
#  facet_wrap(~ Category, scales = "free_y") +
#  theme(legend.position = "bottom") + theme_classic(base_size=6) + 
#  theme(
#    base_size=6,
#    legend.text = element_text(size = 6),
#    legend.position = "None"
#  ) 

ratio_coherent_circa_gg <- ft_coherent_circa_df %>%
  group_by(Category) %>%
  mutate(Ratio = Value / sum(Value)) %>%
  filter(Type == "Coherent") %>% 
  ggplot(aes(x = Category, y = Ratio)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(x = "Category", y = "Proportion of Coherent", title = "Coherent Ratio by Category") +
  theme_classic(base_size = 6)


#ft_change_circa <- fisher_change_test(sim_res_corrected, circannual_gene_ids)
#print(ft_change_circa$fisher) 

#ft_change_circa_df <- ft_change_circa$table %>% as.data.frame() %>% 
#  rownames_to_column("Category") %>% 
#  pivot_longer(c(Change, `No-change`)) %>% 
#  setNames(c("Category", "Type", "Value")) %>% 
#  mutate(Category = ifelse(Category == "Signif.", "Circa." ,Category))


#ft_change_circa_gg <-
#  ggplot(ft_change_circa_df, aes(x = Type, y = Value, fill = Type)) +
#  geom_bar(stat = "identity", position = "dodge") +
#  facet_wrap(~ Category, scales = "free_y") +  # カテゴリごとに個別のグラフ
#  labs(title = "", x = "", y = "# Gene") +
#  scale_fill_manual(values = c("skyblue", "grey")) +
#  theme_minimal() + 
#  facet_wrap(~ Category, scales = "free_y") +
#  theme(legend.position = "bottom") + theme_classic(base_size=6) + 
#  theme(
#    base_size=6,
#    legend.text = element_text(size = 6),
#    legend.position = "None"
#  ) 

#ratio_change_circa_gg <-ft_change_circa_df %>%
#  group_by(Category) %>%
#  mutate(Ratio = Value / sum(Value)) %>%
#  filter(Type == "Change") %>% 
#  ggplot(aes(x = Category, y = Ratio)) +
#  geom_bar(stat = "identity", fill = "steelblue") +
#  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
#  labs(x = "Category", y = "Proportion of Change", title = "Change Ratio by Category") +
#  theme_classic(base_size = 6)


#proportion_gg <- gridExtra::grid.arrange(ft_change_circa_gg, 
#                        ratio_change_circa_gg, 
#                        ft_coherent_circa_gg,
#                        ratio_coherent_circa_gg, nrow=2, widths=c(1.5,1)) 

#proportion_gg <- gridExtra::grid.arrange(ft_coherent_circa_gg, ratio_coherent_circa_gg, nrow=1#, widths=c(1.5,1)) 

#circa_analysis <- gridExtra::grid.arrange(wilcox_ccf_plot_circa, 
#                                          proportion_gg, nrow=2, heights=c(1,1)) 

circa_analysis <- gridExtra::grid.arrange(wilcox_ccf_plot_circa, ratio_coherent_circa_gg, 
                                          nrow=1, widths=c(2,1)) 

ggplot2::ggsave('FiguresV2/Circannual_analysis_v5.pdf',circa_analysis, 
               width=7.13, height =3.0, units = "in")

plot(circa_analysis)

# The limited number of observations below the threshold is an expected outcome, given that a conservative statistical test was employed.
```

# Sex hormone analysis

### Load Environmental data and GSI

Load signal importance generated from \*\*.py. See our method for detail estimation process. Convert twice per month data into once per month to align with the RNA-seq data.

```{r LoadSignalImportance}


Data1input = read.csv("../data/20151001_20171015_env_split.csv", row.names = 1)
Data1output = read.csv("../data/20151001_20171015_out_split.csv", row.names = 1)

sample_point <- format(as.Date(Data1output$Time_raw, format = "%Y-%m-%d %H:%M:%S"), "%y-%m/%d")
Data1output = Data1output %>% mutate(sample_point = sample_point)
GSIData <- Data1output[is.element(sample_point, colnames(medaka_2y_RNAseq_ave)),]

sample_point <- format(as.Date(Data1input$Time_raw, format = "%Y-%m-%d"), "%y-%m/%d")
Data1input = Data1input %>% mutate(sample_point = sample_point)
EnvData <- Data1input[is.element(sample_point, colnames(medaka_2y_RNAseq_ave)),]

head(signal_importance)
head(GSIData)
head(EnvData)
```

### Comparison of sex hormone (estrogen receptor) and Signal importance.

#### 1. Extract hormone genes using the following queries: 'medaka_female_sex_hormone_evidence.html'

```{r echo=FALSE, results='asis'}
htmltools::includeHTML("./medaka_female_sex_hormone_evidence.html")
```

```{r sexhormoneextract}

# Keyword query for the Symbol column (gene names)
fs_hormn_query <- c(
  "cyp19a1a", "cyp19b",
  "cyp17a2", "cyp17", "cyp11a", "cyp11b", "cyp21a2",
  "star", "hsd3b", "hsd17b1", "hsd17b3", "hsd17b4", "hsd17b8", "hsd17b14",
  "srd5a1", "srd5a2", "srd5a3",
  "esr1", "esr2", "pgr", "ar-alpha",
  "nr5a1", "fshr", "lhcgr"
)

# Regex pattern for Nr Description search.
# "star" is excluded to avoid spurious matches (too short/common).
fs_hormn_query_regex <- paste0(
  "\\b(", paste(setdiff(fs_hormn_query, "star"), collapse = "|"), ")\\b"
)

# KEGG Orthology terms for steroidogenesis enzymes
fs_hormn_kos <- c("K07434", "K00513", "K00070", "K12343")

# Filter genes by:
# (1) exact Symbol match
# (2) regex match in Nr Description
# (3) KEGG Orthology term
# Then deduplicate by Symbol and retain only genes present in the expression matrix.
sex_hormone_genes <- gene_profiles %>%
  filter(
    tolower(Symbol) %in% fs_hormn_query |
    str_detect(Nr.Description, regex(fs_hormn_query_regex, ignore_case = TRUE)) |
    Kegg.Orthology %in% fs_hormn_kos
  ) %>%
  distinct(Symbol, .keep_all = TRUE) %>%
  filter(gene_id %in% rownames(medaka_2y_RNAseq_ave))

sex_hormone_genes %>%
  dplyr::select(Symbol, Kegg.Orthology, Nr.Description)
# %>% write.csv("Sex_hormone_genes.csv")
```

#### 2. Visualize the dynamics of sex hormone genes

```{r genevisualize}

fs_plot_df <- medaka_2y_RNAseq_ave_norm[sex_hormone_genes$gene_id,] %>% 
  t() %>% as.data.frame() %>% mutate(sample_point=c(1:24)) %>% 
  pivot_longer(cols=sex_hormone_genes$gene_id) %>% 
  left_join(symbol_to_geneid, by = c("name"="gene_id"))


fsg_dyn_gplot <- ggplot(fs_plot_df, aes(x = sample_point, y = value)) +
  geom_point() + geom_line() + 
  labs(x = "Sample Date", y = "Value", title = "") +
  scale_x_continuous(
      breaks = seq(1, 26, by = 4),  # ラベルを設定する位置
      labels = c("10", "02", "06","10","02","06","10")
    ) + 
    ylab("Normalize") + xlab("Time (month)") + 
  facet_wrap(~Symbol, nrow=4) + theme_classic()
fsg_dyn_gplot
```

#### 2.5. See all comparison.

```{r allcomp}
library(tidyverse)
library(patchwork)
library(cowplot)

# ───────────────────────────── 1. セル描画関数 ────────────────────────────
comp_plot_cell <- function(geneid, comp_value, color, lastid){
  cpl = 
    tibble(sample_point = 1:24,
         gene_z  = scale(as.numeric(medaka_2y_RNAseq_ave_norm[geneid, ])),
         comp_z  = scale(as.numeric(comp_value))) |>
    ggplot() +
      geom_line (aes(sample_point, gene_z , colour = "g"), size = .25) +
      geom_point(aes(sample_point, gene_z , colour = "g"), size = .01) +
      geom_line (aes(sample_point, comp_z , colour = "c"), size = .25) +
      geom_point(aes(sample_point, comp_z , colour = "c"), size = .01) +
      scale_colour_manual(values = c(g = "black", c = color), guide = "none") 
    if(geneid == lastid){
      cpl + theme_classic(base_size=6) + 
        theme(
            axis.line.y = element_blank(),     # y軸の線を非表示
            axis.ticks.y = element_blank(),    # y軸の目盛りを非表示
            axis.ticks.x = element_blank(),    # y軸の目盛りを非表示
            axis.text.y = element_blank(),     # y軸のラベルを非表示
            axis.title.y = element_blank()     # y軸のタイトルも非表示
        ) + 
        xlab("Time (month)") + 
          scale_x_continuous(
            breaks = seq(1, 26, by = 4),  # ラベルを設定する位置
            labels = c("10", "02", "06", "10", "02", "06", "10")
        ) %>% return()
    }else{
      cpl + theme_void(base_size=6) %>% return()
    }
}

# ───────────────────────────── 2. データ準備 ─────────────────────────────
signal_importance_tmp <- signal_importance |>
  set_names(c("Time_raw","W_SR","W_WT","W_DL","W_GSI"))



EnvDataDf <- dplyr::bind_cols(
  dplyr::select(signal_importance_tmp, W_SR, W_WT, W_GSI),
  dplyr::select(EnvData, SR, WT),
  dplyr::select(GSIData, Female_mean, deltaFemaleGSI)
)

env_names <- colnames(EnvDataDf)          # 7 変数
gene_ids  <- sex_hormone_genes$gene_id    # 16 遺伝子
env_cols  <- setNames(c("#EC2526",  "#3B54A4", "#0F8141", 
                        "#F5988F",  "#8DA3E6", "#6ACF9E", "grey"), env_names)

make_cell <- function(g, e){
  comp_plot_cell(g, EnvDataDf[[e]], env_cols[[e]], gene_ids[length(gene_ids)])
}

plot_tbl <- expand_grid(gene = gene_ids, env = env_names) |>   # gene 行優先
  mutate(cell =  purrr::map2(gene, env, make_cell))

# ───────────────────────────── 3. 16×7 格子を生成 ────────────────────────
big_plot <- wrap_plots(plot_tbl$cell,
                       ncol = length(env_names), byrow = TRUE) &
            theme(plot.margin = margin(1, 1, 1, 1, "mm"))


# ───────────────────────────── 4. 行・列ラベルを後付け ───────────────────
# 列（環境）の中心 x 座標
col_pos <- (seq_along(env_names) - .5) / length(env_names)
# 行（遺伝子）の中心 y 座標（上が 1、下が 0）
row_pos <- 1 - (seq_along(gene_ids) - .5) / length(gene_ids)


symbols <- filter(symbol_to_geneid, is.element(gene_id, gene_ids))$Symbol
final_plot_ <- ggdraw(big_plot) +
  theme(plot.margin = margin(t = 5, r = 5, b = 5, l = 5, unit = "mm"))　+
  ## 列ラベル（環境）
  pmap(list(col_pos, env_names),
       ~draw_text(..2, x = ..1, y = 1, size = 7,
                  hjust = 0.5, vjust = 0)) +
  ## 行ラベル（遺伝子）
  pmap(list(row_pos, symbols),
       ~draw_text(..2, x = 0.00, y = ..1, angle = 90, size = 7,
                  hjust = 0.5, vjust = 0.5)) 

final_plot_
```

#### 3. Investigate the correlation distribution

```{r coe}
# correlation
SRimp.hormone.coef <- sapply(sex_hormone_genes$gene_id, function(name){cor(medaka_2y_RNAseq_ave_norm[name,], signal_importance$SR)})
WTimp.hormone.coef <- sapply(sex_hormone_genes$gene_id, function(name){cor(medaka_2y_RNAseq_ave_norm[name,], signal_importance$WT)})
GSIimp.hormone.coef <- sapply(sex_hormone_genes$gene_id, function(name){cor(medaka_2y_RNAseq_ave_norm[name,], signal_importance$GSI)})
SR.hormone.coef <- sapply(sex_hormone_genes$gene_id, function(name){cor(medaka_2y_RNAseq_ave_norm[name,], EnvData$SR)})
WT.hormone.coef <- sapply(sex_hormone_genes$gene_id, function(name){cor(medaka_2y_RNAseq_ave_norm[name,], EnvData$WT)})
GSI.hormone.coef <- sapply(sex_hormone_genes$gene_id, function(name){cor(medaka_2y_RNAseq_ave_norm[name,], GSIData$Female_mean)})
dGSI.hormone.coef <- sapply(sex_hormone_genes$gene_id, function(name){cor(medaka_2y_RNAseq_ave_norm[name,], GSIData$deltaFemaleGSI)})

fsg_hist_df <- tibble(
  W_SR =  SRimp.hormone.coef,
  W_WT =  WTimp.hormone.coef,
  W_GSI = GSIimp.hormone.coef,
  SR    =  SR.hormone.coef,
  WT    =  WT.hormone.coef,
  GSI   =  GSI.hormone.coef,
  dGSI  =  dGSI.hormone.coef
) %>% 
  pivot_longer(cols = everything(),
               names_to = "variable",
               values_to = "coef") 
#%>% 
#  mutate(variable = factor(variable, levels = c(
#    "W_SR", "W_WT", "W_GSI", "SR","WT", "GSI", "dGSI"
#  )))


fsg_hist_plot <- 
  ggplot(fsg_hist_df, aes(x = coef)) +
  geom_histogram(fill = "grey", colour = "black") +
  facet_wrap(~ variable, ncol = 7) +
  scale_x_continuous(limits = c(-1, 1),
                     breaks = seq(-1, 1, 1)) +
  labs(x = "Pearson correlation coefficient",
       y = "Count") +
  theme_classic(base_size = 6) + theme(strip.text =element_blank()) 



#+
final_plot <- final_plot_ + 
  pmap(list(rep(row_pos,each=length(col_pos))+0.01, 
            rep(col_pos, times=length(row_pos))-0.04, paste("r",round(fsg_hist_df$coef,2), sep=": ")),
       ~draw_text(..3, x = ..2, y = ..1,  size = 6,
                  hjust = 0.5, vjust = 0.5)) 

#final_plot 

sex_env_comp_gg <- gridExtra::grid.arrange(
  final_plot, fsg_hist_plot, nrow = 2, heights=c(5,0.7))

#ggsave("FiguresV2/SexHormoneComp2.pdf", final_plot, width=7.08, height=6.69)
```

#### 4. statistical test

```{r loadnullmodel}
# laod null model
data_dir = "./TV0603_alpha0win3noisemean"
SRimp.randomized.coef_ <- read.csv(file.path(data_dir, paste("randomized_SRimp.csv", sep="")), header=FALSE)
WTimp.randomized.coef_ <- read.csv(file.path(data_dir, paste("randomized_WTimp.csv", sep="")), header=FALSE)
GSIimp.randomized.coef_ <- read.csv(file.path(data_dir, paste("randomized_GSIimp.csv", sep="")), header=FALSE)

SR.randomized.coef_ <- read.csv(file.path(data_dir, paste("randomized_SR.csv", sep="")), header=FALSE)
WT.randomized.coef_ <- read.csv(file.path(data_dir, paste("randomized_WT.csv", sep="")), header=FALSE)
GSI.randomized.coef_ <- read.csv(file.path(data_dir, paste("randomized_GSI.csv", sep="")), header=FALSE)
dGSI.randomized.coef_ <- read.csv(file.path(data_dir, paste("randomized_dGSI.csv", sep="")), header=FALSE)


#Align the sample date
sample_point <- format(as.Date(signal_importance_$Time_raw, format = "%Y-%m-%d %H:%M:%S"), "%y-%m/%d")

SRimp.randomized.coef <- SRimp.randomized.coef_[,is.element(sample_point, colnames(medaka_2y_RNAseq_ave))]
WTimp.randomized.coef <- WTimp.randomized.coef_[,is.element(sample_point, colnames(medaka_2y_RNAseq_ave))]
GSIimp.randomized.coef <- GSIimp.randomized.coef_[,is.element(sample_point, colnames(medaka_2y_RNAseq_ave))]
SR.randomized.coef <- SR.randomized.coef_[,is.element(sample_point, colnames(medaka_2y_RNAseq_ave))]
WT.randomized.coef <- WT.randomized.coef_[,is.element(sample_point, colnames(medaka_2y_RNAseq_ave))]
GSI.randomized.coef <- GSI.randomized.coef_[,is.element(sample_point, colnames(medaka_2y_RNAseq_ave))]
dGSI.randomized.coef <- dGSI.randomized.coef_[,is.element(sample_point, colnames(medaka_2y_RNAseq_ave))]
```

```{r surroagate_test_for_sex_hormone_genes}

SRimp_sexhgene_statres <- pbapply::pbsapply(
  sex_hormone_genes$gene_id,
  surrogate_test,
  nullmodel_mat = SRimp.randomized.coef,
  realmodel_vec = as.numeric(signal_importance[['SR']])
)  %>% t() %>% as.data.frame()
colnames(SRimp_sexhgene_statres) <- c("cor", "raw_pval")

WTimp_sexhgene_statres <- pbapply::pbsapply(
  sex_hormone_genes$gene_id,
  surrogate_test,
  nullmodel_mat = WTimp.randomized.coef,
  realmodel_vec = as.numeric(signal_importance[['WT']])
)  %>% t() %>% as.data.frame()
colnames(WTimp_sexhgene_statres) <- c("cor", "raw_pval")

GSIimp_sexhgene_statres <- pbapply::pbsapply(
  sex_hormone_genes$gene_id,
  surrogate_test,
  nullmodel_mat = GSIimp.randomized.coef,
  realmodel_vec = as.numeric(signal_importance[['GSI']])
)  %>% t() %>% as.data.frame()
colnames(GSIimp_sexhgene_statres) <- c("cor", "raw_pval")

SR_sexhgene_statres <- pbapply::pbsapply(
  sex_hormone_genes$gene_id,
  surrogate_test,
  nullmodel_mat = SR.randomized.coef,
  realmodel_vec = as.numeric(EnvData$SR)
)  %>% t() %>% as.data.frame()
colnames(SR_sexhgene_statres) <- c("cor", "raw_pval")

WT_sexhgene_statres <- pbapply::pbsapply(
  sex_hormone_genes$gene_id,
  surrogate_test,
  nullmodel_mat = WT.randomized.coef,
  realmodel_vec = as.numeric(EnvData$WT)
)  %>% t() %>% as.data.frame()
colnames(WT_sexhgene_statres) <- c("cor", "raw_pval")

GSI_sexhgene_statres <- pbapply::pbsapply(
  sex_hormone_genes$gene_id,
  surrogate_test,
  nullmodel_mat = GSI.randomized.coef,
  realmodel_vec = as.numeric(GSIData$Female_mean)
)  %>% t() %>% as.data.frame()
colnames(GSI_sexhgene_statres) <- c("cor", "raw_pval")

dGSI_sexhgene_statres <- pbapply::pbsapply(
  sex_hormone_genes$gene_id,
  surrogate_test,
  nullmodel_mat = dGSI.randomized.coef,
  realmodel_vec = as.numeric(GSIData$deltaFemaleGSI)
)  %>% t() %>% as.data.frame()
colnames(dGSI_sexhgene_statres) <- c("cor", "raw_pval")
  

# result summary
df_list <- list(
  W_SR  = SRimp_sexhgene_statres,
  W_WT  = WTimp_sexhgene_statres,
  W_GSI = GSIimp_sexhgene_statres,
  SR     = SR_sexhgene_statres,
  WT     = WT_sexhgene_statres,
  GSI    = GSI_sexhgene_statres,
  dGSI   = dGSI_sexhgene_statres
)
imap_dfr(df_list, ~{
  .x %>%                                   # 各データフレーム
    mutate(bonfp = p.adjust(raw_pval,
                            method = "bonferroni")) %>%
    summarise(
      dataset = .y,  
      semi_significant = sum(raw_pval < 0.05),# リスト名を保持
      signigicant   = sum(bonfp < 0.1)           # 有意行数
    )
}) %>% gt()

```

#### 4. detail plot

```{r detail}

SR_sexhgene_statres %>% mutate(bonfp = p.adjust(raw_pval, method = "fdr"))
SRimp_sexhgene_statres  %>% mutate(bonfp = p.adjust(raw_pval, method = "fdr"))

GSIimp.signif.geneids <- GSIimp_sexhgene_statres %>%  mutate(bonfp = p.adjust(raw_pval, method = "bonferroni")) %>% filter(bonfp < 0.1) %>% rownames()

WT.signif.geneids <- WT_sexhgene_statres %>%  mutate(bonfp = p.adjust(raw_pval, method = "bonferroni")) %>% filter(bonfp < 0.1) %>% rownames()

symbol_to_geneid %>% filter(is.element(gene_id, GSIimp.signif.geneids)) %>% print()
symbol_to_geneid %>% filter(is.element(gene_id, WT.signif.geneids)) %>% print()
```

### 
